### 2017-12-12 paper/main.tex
Add report to repo

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{pdfpages}
\usepackage{minted}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\graphicspath{ {images/} }
\title{RBE/CS549 Project: Boat Tracking/Detection}
\author{Jordan Burklund \and James Kuszmaul}
\begin{document}
\maketitle

\section{Task}

We are currently working on a robotic sailboat for the
International Robotic Sailing Competition (\url{http://sailbot.org}).

In navigating a sailboat, we would like to be able to
detect and maneuver around various objects
that are floating on the surface of the water. As such,
it is desirable to be able to (a) identify and (b)
predict the motion of objects on the water. Computer
Vision is a potentially convenient way to do this.
We are not specifically aware of any particularly
thorough published work on this particular application
(while we have used vision in the past, it has been
extremely simple detection of bright-colored buoys
on the water).

The goal of this project shall be to use some existing
video to try and detect and track objects in video,
and to understand how well we can develop such a
detector without having a labelled dataset. We will,
for the time being, not worry about being able to run
the algorithm in real time onboard the boat (although
practical considerations mean it must be able to
perform fast enough for us to evaluate it).

\section{Data}

We have approximately 240GB of HD video from a GoPro
mounted on the prow of a large motorboat. While we
do not have the processing power to fully take advantage
of every frame of data, we will pull reasonable samples
of data to work with.  Specifically, we will generate a test dataset of reduced sized that includes the following important characteristics:

\begin{itemize}
\item With/without land (to see different horizon conditions).
\item Varying amounts of other objects/traffic.
\item Areas with significant camera movement in either pitch or roll (may extend to both pitch and roll depending on progress).
\item Different water conditions (starting with flat, calm water, and progressing to more chaotic water conditions as time allows)
\item Objects both near/far away (including near the horizon)
\item Frames that include water droplets on the lens
\end{itemize}

The boat that the data is collected on is relatively stable
and the GoPro is mounted in the same spot for the length
of the video. As such, we should be able to safely ignore
the rotational pose of the boat (when we apply these
algorithms to our final boat, we will have to account for
the much more noticeable movement of that boat).

\begin{figure}[H]
\includegraphics[width=12cm]{sample1}
\centering
\caption{Sample image 1}
\end{figure}
\begin{figure}[H]
\includegraphics[width=12cm]{sample2}
\centering
\caption{Sample image 1}
\end{figure}

\section{General Approach}

Overall, the object pipeline is as follows:

\begin{enumerate}
\item We read a frame of video and convert it to HSV space
\item Perform a slew of static transforms to try and determine,
      based on the content of a single frame, which areas are and
      are not water/sky/land/objects
\item Use optical flow information from a given pair of frames to
      locate the moving objects in a frame
\item Combine the above to determine a set of viable blobs
\item Associate the blobs with each other and with preexisting objects,
      creating new objects for blobs that can't be associated with any
      preexisting objects
\item Update objects in preparation for the next frame, updating
      positions using velocities, reducing and increasing confidence
      scores as appropriate, deleting objects with sufficiently low
      scores
\end{enumerate}

\section{Image Segmentation}

\section{Optical Flow}

For the underlying optical flow algorithm, we use Farneback's
[INSERT CITATION HERE] which is accessible in matlab as...

\section{Evaluation Section from Proposal}

Because we do not have a labeled data set, we must develop
some way evaluate the performance of different algorithms.

In order to do this, we propose two general cost functions.
The first will evaluate the motion tracking features.
Essentially, when motion tracking, our objective is to
correctly predict the future path of an object. As such,
the cost function will assume a sane object detector and
use the difference between predicted frame-to-frame position
and actual frame-to-frame position (from the object detector)
as the cost.

For the object detector itself, a bit more complication is
required, and we hypothesize that the cost function will
consist of the following terms:

\begin{description}
\item[Object Deletion] Any unexpected object deletion shall
   be penalized, in order to avoid constantly losing track
   of objects between frames or over/under segmentation of
   individual objects. ``Expected'' object deletion would
   occur when an object that is already moving out of frame
   or behind another object disappears. In theory we also
   care about excessive object \emph{creation}; however,
   because the vast majority of created objects are also
   deleted eventually, the additional complication should
   be unnecessary. There is an open question about how
   to treat the deletion of objects that disappear over
   the horizon; however, such occurrences are hopefully
   rare enough or easily detectable enough that their
   cost should be insignificant (we don't want to
   accidentally \emph{not} penalize flickering objects
   that happen to fall on the horizon).
\item[Number of Objects] In order to avoid either a
   detector that never deletes objects or one that
   divides real objects into far too many individual
   objects, we shall penalize the number of objects
   identified in any given frame. This may be quadratic
   or some higher order function in order to avoid
   penalizing small numbers of objects excessively.
\item[Unaccounted for Area] To avoid an object
   detector that simply never returns any objects,
   we shall compare the area accounted for by the
   identified objects and the area that ``should''
   contain objects (per some naive estimate), and
   penalize any unaccounted for area.
   As an initial method for identifying this area,
   we propose that we (a) develop some simple detector
   for identifying sea water (it's mostly blue...),
   and (b) make a simple assumption about the location
   of the horizon (the existing video is relatively stable,
   so the horizon should be, and this way we won't have
   to worry about the sky). Any area below the estimated
   horizon that is not sea water ``should'' be part of
   an object (this includes the shore).
\end{description}

Obviously, the above cost function will not provide a perfect
measure of algorithm performance. However, between it and
qualitative sanity checks, we should be able to provide
a good estimate of the performance of our algorithms.

\section{Responsibilities Section from Proposal}
\begin{description}
\item[James Kuszmaul] Work on basic object detection and how
  to retain object associations between frames/measure .

\item[Jordan Burklund] Develop methods to segment and mask pixels that represent water. Explore methods using contours for object detection.
\end{description}
\end{document}

### 2017-12-12 paper/main.tex
Add citation

@@
 For the underlying optical flow algorithm, we use Farneback's
-[INSERT CITATION HERE] which is accessible in matlab as...
+\cite{farneback2003} which is accessible in matlab as
+\texttt{opticalFlowFarneback}.
+
+As an overview, the way that this optical flow algorithm works is to
 
@@
 \end{description}
+
+\bibliographystyle{ieeetr}
+\bibliography{cite}
+
 \end{document}

### 2017-12-12 paper/main.tex
Worked on optical flow section

@@
 \usepackage{minted}
+\usepackage{amsmath}
+\usepackage{bm}
 \usepackage{hyperref}
@@
 
+\subsection{Underlying Algorithm}
+
 For the underlying optical flow algorithm, we use Farneback's
@@
 As an overview, the way that this optical flow algorithm works is to
+approximate the neighborhood of every pixel as some quadratic function
+$f(\mathbf{x}) = \mathbf{x}^T \mathbf{A} \mathbf{x} + \mathbf{b}^T \mathbf{x} + c$
+where $\mathbf{x}$ is the location around any given pixel. By estimating these
+polynomials about any given pixel between two frames, we can try to estimate how
+much the image has moved in the neighborhood of that pixel. Actual
+implementations of the algorithm contain some additional filtering/smoothing and
+function on multiple levels of an image pyramid to estimate optical flow at
+different scales. You can also adjust how much of the neighborhood around a
+pixel is used to estimate the polynomial, and various other optimizations.
+
+In practice, we just used the default options to the matlab method (which uses
+a 3-level pyramid scaling image size by a factor of 2, with a 5-pixel
+neighborhood size and a Gaussian filter on the output of size 15x15), as these
+options tended to produce good results.
+
+\subsection{Expected Flow}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{example_flow.png}
+\caption{Optical Flow output (shown only for every few pixels) between two
+frames}
+\label{fig:example_flow}
+\end{figure}
+
+In Figure \ref{fig:example_flow} we show the output of
+\texttt{opticalFlowFarneback} on a given pair of frames. The most immediate need
+in this image is to try and account for the field of flow produced along the
+water. This corresponds to the real movement of the plane of the water relative
+to the boat. We can attempt to model this with the following assumptions:
+\begin{itemize}
+\item That the camera is pointed straight and level
+\item There is no substantial roll or yaw angular rates of turn occurring
+\item The camera is an ideal pinhole camera with no substantial lens distortions
+      or the such
+\item There is no sideways relative velocity between the boat and the water
+\end{itemize}
+
+We can define the following variables:
+
+\begin{tabular}{c|c}
+Name & Description \\ \hline
+$h$ & Height of the camera above the water \\
+$v_h$ & The current upwards velocity of the camera \\
+$v_b$ & Forwards velocity of the boat relative to the water \\
+$x_b, y_b$ & Position coordinates in the boat frame \\
+$x_f, y_f$ & X and Y position in the 2-D image frame \\
+$\dot{x}_f, \dot{y}_f$ & The expected optical flow (derivatives of $x_f, y_f$)
+\\
+$K$ & Some constant describing the field of view of the camera \\
+\end{tabular}
+
+The boat frame has an origin immediately below the camera on the surface of
+the water, is attached to the boat, has an x-axis pointing straight forwards,
+a y-axis pointing straight port (left in the camera view) and a z-axis pointing
+straight upwards, with the camera at position $(0, 0, h)$ in this frame.
+
+The image frame is scaled so that the center of the image is the origin and
+the image is exactly 1 unit tall, with the x-axis being the side-to-side axis
+and the y-axis being up-and-down.
+
+Strictly speaking $K = 2\sin\frac\theta2$ if the total vertical field-of-view
+of the camera is $\theta$, but we end up estimating it anyways.
+
+These give us:
+
+\begin{align*}
+x_f =& \frac{-y_b}{Kx_b} \\
+y_f =& \frac{-h}{Kx_b} \\
+\dot{x}_f =& \frac{-y_bv_b}{Kx^2_b} \\
+\dot{y}_f =& -\frac{1}{Kx_b}(\frac{hv_b}{x_b} + v_h) \\
+\end{align*}
+
+And solving for $x_b, y_b$ in terms of $x_f, y_f$ and then substituting in to
+give us the expected optical flow at any given image coordinate $x_f, y_f$ we
+get:
+
+\begin{align}
+\dot{x}_f =& \frac{-v_bKx_fy_f}{h} \\
+\dot{y}_f =& \frac{v_hy_f - Kv_by_f^2}{h}
+\end{align}
+
+If we ignore the actual values of the constants and just turn them into
+something to be estimated, while also adding in a parameter where we
+assume that $y_f$ may not be perfectly centered, we can turn this into a
+linear parameter estimation problem as:
+
+\begin{align}
+\dot{x}_f =& ax_fy_f + bx_f \\
+\dot{y}_f =& cy_f^2 + dy_f + e
+\end{align}
+
+To actually perform the least-squares fitting we use \texttt{fitlm} in
+matlab. Normally, this just uses a normal least-squares estimation
+(similar to doing \texttt{A \ b}), but adds some additional
+options. We tried to make use of these options, including a outlier
+rejection algorithm, but the effects were minimal and substantially
+increased the processing time required to execute the code.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{diff_magnitude}
+\caption{Difference between the measured and modelled optical flow.
+         Dark blue means perfect match, bright yellow means a poor match.}
+\label{fig:diff_magnitude}
+\end{figure}
+
+When we compare the expected magnitude of the optical flow to the
+actual magnitude of the optical flow in Figure \ref{fig:diff_magnitude}
+(this is the same image as Figure \ref{fig:example_flow}),
+in which you can see a few features:
+\begin{itemize}
+\item The boat in the left half of the image shows up as a cyan blob, distinct
+      from the water around it.
+\item The curved metal in the bottom left screws with the optical flow
+      algorithm, causing substantial variation from the model.
+\item In the bottom right, due to either lens distortion or hard to measure
+      waves, the optical flow varies substantially from the model.
+\end{itemize}
+
+As a result of the poor results in the very bottom of the image and in the
+area around the metal bar, we completely ignore and filter out these regions,
+as they are relatively unlikely to contain obstacles (or if they did, our
+boat would be running them over anyways).
+
+We also note that in general objects appearing more distantly in the center of
+the image will be hard to identify, as the absolute value of the difference
+in magnitude from the modelled magnitude will be low (because we are looking
+at optical flow in the image, not real life, so far away objects have
+low velocities in the image), and so simple thresholds will not work well.
+We experimented with reducing the required threshold in the center of the image,
+which did improve tracking.
+
+It is also the case that any stationary, low-lying obstacles (e.g. buoys)
+will be indistinguishable from the water and so will not be detected
+by this half of the filtering.
+
+\section{Blob Detection and Filtering}
+
+\section{Object Matching}
+
+\section{Results}
+
+\section{Future Results}
 

### 2017-12-13 paper/main.tex
Fix package includes
Add Pixel Level Water Segmentation section

@@
 \usepackage{pdfpages}
-\usepackage{minted}
 \usepackage{amsmath}
@@
 \usepackage[margin=1in]{geometry}
+\usepackage{float}
+\usepackage{chngcntr}
+\counterwithin{figure}{section}
 \graphicspath{ {images/} }
@@
 Overall, the object pipeline is as follows:
+\begin{figure}[H]
+\includegraphics[width=16cm]{algorithm_flowchart}
+\centering
+\caption{High-Level Image Processing Pipeline}
+\end{figure}
 
@@
 
-\section{Image Segmentation}
+\section{Pixel Level Water Segmentation}
+In the first part of the pipeline, we look at individual images and attempt to provide a pixel level detection of potential objects to track.
+This step of the pipeline provides a confidence value for each pixel in the frame for confidence that the pixels is an object to track.
+The image is converted to HSV color space from the original RGB colorspace since this gives us better segmentation results.
+The HSV space has better color properties for segmentation since the water is typical near one particular color, but can have variations in brightness.
+This also allows the segmentation to easily segment non-water pixels by the hue value, since the water tends to be near the blue and green hue values in our test images.
+
+\subsection{Naiive K-Means Classifier}
+We initially started with a basic two-class k-means classifier in order to observe how well it would perform.
+The classifier was trained on the first image in the video sequence, and the l2 distance to the two cluster centers was used to classify each pixel in subsequent frames.
+Surprisingly enough, the 2-class k-means classifier worked relatively robustly on the videos that we provided the algorithm, as show in figure \ref{fig:kmeans}.
+The water pixels are mostly labeled as one class, and pixels that not water are mostly labeled as another class with a little bit of misclassification noise.
+The boat and it's wake are clearly labeled as potential obstacles to track, which is what we want.
+The land areas are labeled as water pixels, but for our implementation we want to ignore tracking the land anyways.
+There is a slight bit of noise in the water reflections towards the right side of figure \ref{fig:kmeans}, but this naiive method still provides very useful information.
+\\
+\newline
+Kmeans does not always associate the same class identifier with the water and non-water pixels each time, so we must perform an intermediary step to detect which class identifier it associates with water pixels.
+To do so, we find the number of pixels in each class in the bottom half of the image, and use the sums to vote for which class is associated with water pixels.
+This assumes that the boat is upright, and that a majority of the pixels in the bottom half of the image are water, and holds true for all of the cases we observed.
+If the water class ID is not what was expected, the cluster centers are swapped so that the water class ID is always a known value.
+
+\begin{figure}
+\label{fig:kmeans}
+\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
+\includegraphics[width=8.5cm]{hsv_kmeans2_result}
+\centering
+\caption{Naiive K-Means Classifier Results}
+\end{figure}
+
+\subsection{Wave Reflection Suppression}
+Although there is some noise in the classification results, the k-means classifier does a decent job of selecting pixels that are areas of water.
+In order to suppress false detections as seen in the right side of the classifier results in figure \ref{fig:kmeans}, we look at the distance values between the pixel's HSV value and the water pixel cluster center to estimate how similar the pixel is to the water class.
+The wave reflections observed in figure \ref{fig:kmeans} tend to very close to the cluster center while true obstacle pixels tend to be further away from the cluster center.
+By using this measure as a confidence value, we can suppress regions that are more likely to be water, and emphasize regions that are more likely to be objects to track.
+Using the distance metric directly provided decent results, but squaring the distance metric value provided significantly better suppression of water and emphasis of potential objects.
+The squared value was then used as a direct measure of confidence that each pixel was a potential object to track.
+
+\subsection{Sky Suppression}
+The k-means classifier provides information about which pixels are water and which pixels are not water, but we would also like the algorithm to only focus on potential obstacles in the water.
+The region of the sky shown in figure \ref{fig:kmeans} is classified as a non-water pixel, and we would like to suppress this response to avoid detecting the sky as an obstacle.
+With only the wave reflection suppression confidences, most of the sky is suppressed, but tends to have higher confidences near land as shown in figure \ref{fig:skysuppress}.
+Since the sky pixels are all connected, and define a very large region, we used the Matlab bwareaopen function that removes binary pixels in connected areas that are smaller than a particular size.
+By looking only at pixels that are classfied as "not water", we can detect the sky region by filtering for pixel clusters that are larger than 1/4 of the area of the image, and set the confidence values for those pixels to zero to suppress them.
+In some cases, potential objects to track were touching the sky region and the bwareaopen results tended to remove those object to track.
+To fix this, an erode operation is done on the non-water class pixels to better separate the sky and potential objects to track.
+
+\begin{figure}
+\label{fig:skysuppress}
+\includegraphics[height=5cm]{hsv_confidence}
+\includegraphics[height=5cm]{hsv_kmeans2_suppressed}
+\centering
+\caption{Sky Supression Before and After}
+\end{figure}
+
+\subsection{Confidence Image Results}
+When the above confidence values and suppressions described above are combined, a final confidence value image is produced.
+Pixels that are more likely to be objects to track have higher confidence values that pixels that are more likely to be water areas.
+Figure \ref{fig:boatconf} shows the detector correctly detecting a boat in the image.
+Figure \ref{fig:towerconf} shows the detector correctly detecting a tower in the ditsance.
+Figure \ref{fig:buoyconf} shows the detector identifying a buoy off in the distance.
+
+\begin{figure}[H]
+\label{fig:boatconf}
+\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
+\includegraphics[width=8.5cm]{hsv_kmeans2_suppressed}
+\centering
+\caption{Boat Confidence Results}
+\end{figure}
+
+\begin{figure}[H]
+\label{fig:towerconf}
+\includegraphics[width=7.9cm]{hsv_kmeans2_tower}
+\includegraphics[width=8.3cm]{hsv_kmeans2_tower_result}
+\centering
+\caption{Tower Confidence Results}
+\end{figure}
+
+\begin{figure}[H]
+\label{fig:buoyconf}
+\includegraphics[width=7.8cm]{hsv_kmeans2_buoy}
+\includegraphics[width=8.5cm]{hsv_kmeans2_buoy_result}
+\centering
+\caption{Buoy Confidence Results}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering

### 2017-12-13 paper/main.tex
Fix spelling errors, make some lines 80 characters

@@
 
-\subsection{Naiive K-Means Classifier}
-We initially started with a basic two-class k-means classifier in order to observe how well it would perform.
-The classifier was trained on the first image in the video sequence, and the l2 distance to the two cluster centers was used to classify each pixel in subsequent frames.
-Surprisingly enough, the 2-class k-means classifier worked relatively robustly on the videos that we provided the algorithm, as show in figure \ref{fig:kmeans}.
-The water pixels are mostly labeled as one class, and pixels that not water are mostly labeled as another class with a little bit of misclassification noise.
-The boat and it's wake are clearly labeled as potential obstacles to track, which is what we want.
-The land areas are labeled as water pixels, but for our implementation we want to ignore tracking the land anyways.
-There is a slight bit of noise in the water reflections towards the right side of figure \ref{fig:kmeans}, but this naiive method still provides very useful information.
-\\
-\newline
-Kmeans does not always associate the same class identifier with the water and non-water pixels each time, so we must perform an intermediary step to detect which class identifier it associates with water pixels.
-To do so, we find the number of pixels in each class in the bottom half of the image, and use the sums to vote for which class is associated with water pixels.
-This assumes that the boat is upright, and that a majority of the pixels in the bottom half of the image are water, and holds true for all of the cases we observed.
-If the water class ID is not what was expected, the cluster centers are swapped so that the water class ID is always a known value.
+\subsection{Naive K-Means Classifier}
+We initially started with a basic two-class k-means classifier in order to
+observe how well it would perform.  The classifier was trained on the first
+image in the video sequence, and the l2 distance to the two cluster centers was
+used to classify each pixel in subsequent frames.  Surprisingly enough, the
+2-class k-means classifier worked relatively robustly on the videos that we
+provided the algorithm, as show in figure \ref{fig:kmeans}.  The water pixels
+are mostly labeled as one class, and pixels that not water are mostly labeled as
+another class with a little bit of misclassification noise.  The boat and it's
+wake are clearly labeled as potential obstacles to track, which is what we want.
+The land areas are labeled as water pixels, but for our implementation we want
+to ignore tracking the land anyways. There is a slight bit of noise in the water
+reflections towards the right side of figure \ref{fig:kmeans}, but this naive
+method still provides very useful information.
+
+Kmeans does not always associate the same class identifier with the water and
+non-water pixels each time, so we must perform an intermediary step to detect
+which class identifier it associates with water pixels.  To do so, we find the
+number of pixels in each class in the bottom half of the image, and use the sums
+to vote for which class is associated with water pixels. This assumes that the
+boat is upright, and that a majority of the pixels in the bottom half of the
+image are water, and holds true for all of the cases we observed.  If the water
+class ID is not what was expected, the cluster centers are swapped so that the
+water class ID is always a known value.
 
@@
 \centering
-\caption{Naiive K-Means Classifier Results}
+\caption{Naive K-Means Classifier Results}
 \end{figure}
@@
 \subsection{Wave Reflection Suppression}
-Although there is some noise in the classification results, the k-means classifier does a decent job of selecting pixels that are areas of water.
-In order to suppress false detections as seen in the right side of the classifier results in figure \ref{fig:kmeans}, we look at the distance values between the pixel's HSV value and the water pixel cluster center to estimate how similar the pixel is to the water class.
-The wave reflections observed in figure \ref{fig:kmeans} tend to very close to the cluster center while true obstacle pixels tend to be further away from the cluster center.
-By using this measure as a confidence value, we can suppress regions that are more likely to be water, and emphasize regions that are more likely to be objects to track.
-Using the distance metric directly provided decent results, but squaring the distance metric value provided significantly better suppression of water and emphasis of potential objects.
-The squared value was then used as a direct measure of confidence that each pixel was a potential object to track.
+Although there is some noise in the classification results, the k-means
+classifier does a decent job of selecting pixels that are areas of water. In
+order to suppress false detections as seen in the right side of the classifier
+results in figure \ref{fig:kmeans}, we look at the distance values between the
+pixel's HSV value and the water pixel cluster center to estimate how similar the
+pixel is to the water class.  The wave reflections observed in figure
+\ref{fig:kmeans} tend to very close to the cluster center while true obstacle
+pixels tend to be further away from the cluster center.  By using this measure
+as a confidence value, we can suppress regions that are more likely to be water,
+and emphasize regions that are more likely to be objects to track.  Using the
+distance metric directly provided decent results, but squaring the distance
+metric value provided significantly better suppression of water and emphasis of
+potential objects.  The squared value was then used as a direct measure of
+confidence that each pixel was a potential object to track.
 
 \subsection{Sky Suppression}
-The k-means classifier provides information about which pixels are water and which pixels are not water, but we would also like the algorithm to only focus on potential obstacles in the water.
-The region of the sky shown in figure \ref{fig:kmeans} is classified as a non-water pixel, and we would like to suppress this response to avoid detecting the sky as an obstacle.
-With only the wave reflection suppression confidences, most of the sky is suppressed, but tends to have higher confidences near land as shown in figure \ref{fig:skysuppress}.
-Since the sky pixels are all connected, and define a very large region, we used the Matlab bwareaopen function that removes binary pixels in connected areas that are smaller than a particular size.
-By looking only at pixels that are classfied as "not water", we can detect the sky region by filtering for pixel clusters that are larger than 1/4 of the area of the image, and set the confidence values for those pixels to zero to suppress them.
-In some cases, potential objects to track were touching the sky region and the bwareaopen results tended to remove those object to track.
-To fix this, an erode operation is done on the non-water class pixels to better separate the sky and potential objects to track.
+The k-means classifier provides information about which pixels are water and
+which pixels are not water, but we would also like the algorithm to only focus
+on potential obstacles in the water.  The region of the sky shown in figure
+\ref{fig:kmeans} is classified as a non-water pixel, and we would like to
+suppress this response to avoid detecting the sky as an obstacle. With only the
+wave reflection suppression confidences, most of the sky is suppressed, but
+tends to have higher confidences near land as shown in figure
+\ref{fig:skysuppress}.  Since the sky pixels are all connected, and define a
+very large region, we used the Matlab bwareaopen function that removes binary
+pixels in connected areas that are smaller than a particular size.  By looking
+only at pixels that are classified as "not water", we can detect the sky region
+by filtering for pixel clusters that are larger than 1/4 of the area of the
+image, and set the confidence values for those pixels to zero to suppress them.
+In some cases, potential objects to track were touching the sky region and the
+bwareaopen results tended to remove those object to track.  To fix this, an
+erode operation is done on the non-water class pixels to better separate the sky
+and potential objects to track.
 
@@
 \centering
-\caption{Sky Supression Before and After}
+\caption{Sky Suppression Before and After}
 \end{figure}
@@
 \subsection{Confidence Image Results}
-When the above confidence values and suppressions described above are combined, a final confidence value image is produced.
-Pixels that are more likely to be objects to track have higher confidence values that pixels that are more likely to be water areas.
-Figure \ref{fig:boatconf} shows the detector correctly detecting a boat in the image.
-Figure \ref{fig:towerconf} shows the detector correctly detecting a tower in the ditsance.
-Figure \ref{fig:buoyconf} shows the detector identifying a buoy off in the distance.
+When the above confidence values and suppressions described above are combined,
+a final confidence value image is produced. Pixels that are more likely to be
+objects to track have higher confidence values that pixels that are more likely
+to be water areas. Figure \ref{fig:boatconf} shows the detector correctly
+detecting a boat in the image.  Figure \ref{fig:towerconf} shows the detector
+correctly detecting a tower in the distance.  Figure \ref{fig:buoyconf} shows
+the detector identifying a buoy off in the distance.
 
@@
 
-\section{Future Results}
+\section{Future Work}
 

### 2017-12-14 paper/main.tex
Fix more 80-character lines
Write a bit on the report

@@
 the much more noticeable movement of that boat).
-
-\begin{figure}[H]
-\includegraphics[width=12cm]{sample1}
-\centering
-\caption{Sample image 1}
-\end{figure}
 \begin{figure}[H]
@@
 \section{Pixel Level Water Segmentation}
-In the first part of the pipeline, we look at individual images and attempt to provide a pixel level detection of potential objects to track.
-This step of the pipeline provides a confidence value for each pixel in the frame for confidence that the pixels is an object to track.
-The image is converted to HSV color space from the original RGB colorspace since this gives us better segmentation results.
-The HSV space has better color properties for segmentation since the water is typical near one particular color, but can have variations in brightness.
-This also allows the segmentation to easily segment non-water pixels by the hue value, since the water tends to be near the blue and green hue values in our test images.
+In the first part of the pipeline, we look at individual images and attempt to
+provide a pixel level detection of potential objects to track.  This step of the
+pipeline provides a confidence value for each pixel in the frame for confidence
+that the pixels is an object to track.
+The image is converted to HSV color space from the original RGB colorspace since
+this gives us better segmentation results. The HSV space has better color
+properties for segmentation since the water is typical near one particular
+color, but can have variations in brightness.  This also allows the segmentation
+to easily segment non-water pixels by the hue value, since the water tends to be
+near the blue and green hue values in our test images.
 
@@
 observe how well it would perform.  The classifier was trained on the first
-image in the video sequence, and the l2 distance to the two cluster centers was
+image in the video sequence, and the L2 distance to the two cluster centers was
 used to classify each pixel in subsequent frames.  Surprisingly enough, the
@@
 
-Kmeans does not always associate the same class identifier with the water and
+K-means does not always associate the same class identifier with the water and
 non-water pixels each time, so we must perform an intermediary step to detect
@@
 \begin{figure}
-\label{fig:kmeans}
 \includegraphics[width=7.8cm]{hsv_kmeans2_orig}
@@
 \caption{Naive K-Means Classifier Results}
+\label{fig:kmeans}
 \end{figure}
@@
 \ref{fig:skysuppress}.  Since the sky pixels are all connected, and define a
-very large region, we used the Matlab bwareaopen function that removes binary
+very large region, we used the Matlab \texttt{bwareaopen} function that removes binary
 pixels in connected areas that are smaller than a particular size.  By looking
@@
 In some cases, potential objects to track were touching the sky region and the
-bwareaopen results tended to remove those object to track.  To fix this, an
+\texttt{bwareaopen} results tended to remove those object to track.  To fix this, an
 erode operation is done on the non-water class pixels to better separate the sky
@@
 \begin{figure}
-\label{fig:skysuppress}
 \includegraphics[height=5cm]{hsv_confidence}
@@
 \caption{Sky Suppression Before and After}
+\label{fig:skysuppress}
 \end{figure}
@@
 a final confidence value image is produced. Pixels that are more likely to be
-objects to track have higher confidence values that pixels that are more likely
+objects to track have higher confidence values than pixels that are more likely
 to be water areas. Figure \ref{fig:boatconf} shows the detector correctly
@@
 \begin{figure}[H]
-\label{fig:boatconf}
 \includegraphics[width=7.8cm]{hsv_kmeans2_orig}
@@
 \caption{Boat Confidence Results}
+\label{fig:boatconf}
 \end{figure}
@@
 \begin{figure}[H]
-\label{fig:towerconf}
 \includegraphics[width=7.9cm]{hsv_kmeans2_tower}
@@
 \caption{Tower Confidence Results}
+\label{fig:towerconf}
 \end{figure}
@@
 \begin{figure}[H]
-\label{fig:buoyconf}
 \includegraphics[width=7.8cm]{hsv_kmeans2_buoy}
@@
 \caption{Buoy Confidence Results}
+\label{fig:buoyconf}
 \end{figure}
@@
 
-\subsection{Expected Flow}
+\subsection{Expected Flow} \label{expected_flow}
 
@@
 
-\section{Blob Detection and Filtering}
+\section{Object Creation and Management}
 
-\section{Object Matching}
+\subsection{Blob Detection and Filtering}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{threshold_bwlabel}
+\caption{Result of thresholding after dilation/labelling}
+\label{fig:threshold_bwlabel}
+\end{figure}
+
+From the results in Figures \ref{fig:boatconf} and
+\ref{fig:diff_magnitude}, we have a set of confidence values that can be used to
+identify where individual obstacles may be. In order to combine the two, we
+simply scale each result by some constant and sum the two to provide a
+detection score. Given this score, we then:
+\begin{enumerate}
+\item Threshold against some constant (tunable) value
+\item Do an erode operation using a circular erosion pattern
+\item Do a dilate in reverse, using a radius roughly twice that of the erosion,
+      so that not only is lost space taken up but nearby segments are connected.
+\item Label the connected regions using \texttt{bwlabel} (corresponding with
+      Figure \ref{fig:threshold_bwlabel}).
+\item For each of these regions, throw out any regions that meet any of the
+      following conditions (resulting in Figure \ref{fig:filtered_labels}:
+  \begin{itemize}
+    \item The object is too ``blue'' (by comparing to some hue-based value,
+          in order to try and filter out any water that got detected)
+    \item The object is too small (suggesting it may not really exist)
+    \item The object is too large (sometimes large (~half the image) connected
+          segments will appear that are virtually never real objects)
+    \item It is moving too slowly relative to us (relative to the camera,
+          the only things that should be that constant are on our
+          boat---anything else should at least contain some noise)
+  \end{itemize}
+\end{enumerate}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{filtered_labels}
+\caption{Result of filtering the blobs in Figure \ref{fig:threshold_bwlabel},
+         with the boat and parts of its wake remaining}
+\label{fig:filtered_labels}
+\end{figure}
+
+Following these steps, we initialize each remaining blob as an object, reducing
+it to a bounding box (rather than an exact set of pixels), a median velocity
+(from the optical flow), a position (the center of the bounding box), a label
+(arbitrarily assigned), and a confidence score (initialized to 0.2, to represent
+an object that we are not entirely sure exists yet).
+
+We then go through and, for each object, attempt to both associate it with other
+newly created objects (in case any objects were over-segmented) as well as
+objects that we've kept track of from the previous frame.
+
+\subsection{Object Matching}
+
+When attempting to match objects, we first update all the old objects so that
+their positions have been shifted by their velocities, to better match objects
+that are moving between frames. Furthermore, we reduce the bounding box sizes on
+old objects to reflect a greater uncertainty; if the object really does exist
+over the entire previous bounding box, then that should be reaffirmed by any
+blobs detected in the current frame.
+
+We then perform the following steps:
+
+\begin{enumerate}
+\item For each pair of new objects, as well as each pairing of one new and one
+      old object, we compute a difference score that is the sum of:
+  \begin{itemize}
+  \item The difference in velocity direction
+  \item One minus the quotient in velocity magnitudes (speed), such that the
+        larger of the two speeds shall always be in the denominator
+  \item The difference in the positions (L2-norm) of the two objects, scaled
+        along each axis by the mean width and mean height between the two
+        objects (such that if the objects are perfectly aligned in one dimension
+        and touching edges in the other, they will be a distance of 1.0 apart).
+  \end{itemize}
+\item We threshold every difference, leaving a matrix of object ``neighbors''.
+\item Iterating over the newly created objects (in arbitrary order), we
+      combine that object with it's immediate neighbors, and then remove
+      all the objects that were combined from future
+      consideration\footnote{Ideally, we might do some sort of search where we
+      combine neighbors-of-neighbors and so on, but for simplicities' sake, and
+      to avoid situations where we end up combining large numbers of objects
+      that all are tenuously connected, we just explore one step out from any
+      given object.}.
+\item In order to combine each set of objects that we associated, we:
+  \begin{itemize}
+    \item Create a bounding box that surrounds all of the combined objects
+    \item Give the amalgamation the label (and time-of-existence) of the oldest
+          object being included, on the assumption that we are interested
+          in keeping tracking going across the longest period of time possible.
+    \item Perform a weighted average of the velocities, weighting by object area.
+    \item Take the maximum of the confidence scores for the individual objects,
+          assign it to the result object, and then increment it up slightly on
+          the assumption that any time we are combining more than one object,
+          we are probably becoming more certain in the overall object's
+          existence.
+  \end{itemize}
+\item Any old objects that were not associated with blobs in the current frame
+      have their confidence scores reduced and, if they dip below a confidence
+      of 0, are deleted permanently.
+\item Finally, for every object being tracked, we keep a history of its last 20
+      positions. Once an object accumulates a history of more than around 5
+      frames (to reduce potential for individual frames causing too much noise),
+      we start to compare the object's overall average position change to its
+      velocity. If the two differ by more than a particular threshold, then we
+      reduce the confidence score again. This helps to mitigate object resulting
+      from the wake of the boat, as such objects generally have optical flows
+      inconsistent with where the object actually moves frame-to-frame.
+\end{enumerate}
+
+\section{Objective Evaluation}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{example_detection}
+\caption{Sample frame of the whole stack running}
+\label{fig:example_detection}
+\end{figure}
+
+In Figure \ref{fig:example_detection}, we see an individual frame from the
+algorithm, where we have:
+\begin{itemize}
+\item The boat identified by a yellow (indicating that it has existed for at
+      least 10 frames) solid (indicating high confidence) box with an arrow
+      pointing leftwards, roughly in the direction of its movement.
+\item Several solid and dashed boxes detecting the wake of the boat.
+\end{itemize}
+
+Longer videos can be found in our presentation:
+
+\begin{description}
+\item[\url{https://drive.google.com/file/d/1FoKWkp5xDSKwp5YyluKrclo_kGFU4gsg/view?usp=sharing}]
+  In this video we have a longer view of the same boat as Figure
+  \ref{fig:example_detection}
+\item[\url{https://drive.google.com/file/d/1ndVSwF77_4va6U8rWT4FGJOXq1AjKL_Z/view?usp=sharing}]
+  This has a slightly shorter capture of a different boat, indicating some
+  degree of robustness to different individual instances
+  \ref{fig:example_detection}
+\end{description}
+
+\subsection{Timing}
+
+As we are interested in ultimately implementing some of these algorithms on our
+actual robot, we are interested in the run time of our algorithms. These
+are extremely rough numbers, from running only single-threaded on a laptop CPU,
+but were reasonably consistent across runs:
+\begin{itemize}
+\item Around 3 seconds total per frame
+\item Around 1 second to call \texttt{opticalFlowFarneback} on the full
+      1920x1080 image
+\item Around half a second to perform the linear fit described in section
+      \ref{expected_flow}
+\item Around a second and a half spread out among other steps
+\end{itemize}
+
+The linear fit step in particular initially took up a substantial amount of
+time, but we discovered that simply subsampling the data and only using around
+10 percent of it worked almost as well while reducing the runtime to the half
+second seen here.
+
+The optical flow timing is exacerbated by the fact that we are running optical
+flow on the entire, reasonably high-resolution, image. Were we to focus better
+on interest regions, or simply use a lower resolution, we could likely achieve
+substantially better results.
+
+\subsection{Robustness}
+
+\begin{figure}[H]
+\includegraphics[width=12cm]{sample1}
+\centering
+\caption{Mirror-like water}
+\label{fig:mirror}
+\end{figure}
+
+While the algorithms do work reasonably well for footage from these particular
+conditions, we had troubles processing some parts of the data. In particular,
+early in certain mornings when not only was the sun interfering with the camera,
+but also in situations such as Figure \ref{fig:mirror}, in which the water is
+relatively flat (reducing texture) and highly reflective, meaning that it is no
+longer particularly blue.
+
 \section{Future Work}
+
+For future work, we need to actually improve the timing (as mentioned above) to
+work in real time, improve robustness for a greater variety of conditions, and
+to increase the generality to function on a less stable platform--our boat will
+certainly be pitching and rolling substantially more than the platform we had
+for this footage.
 

### 2017-12-14 paper/main.tex
Add section on cost function, remove unused proposal sections

@@
 \maketitle
+
+\tableofcontents
 
@@
 
+While we did not end up making much use of it, we did attempt to write a cost
+function that would offer more objective evaluations of our algorithms (as we do
+not have labeled data to perform some simple check against). The
+motivation for this was both to (a) be able to run our algorithms over larger
+segments of data without manually scoring it and (b) help us to better
+understand our own priorities when writing our detection algorithms.
+
+The cost function ultimately consisted of:
+
+\begin{itemize}
+\item A term penalizing the current number of objects (to avoid bloating the
+      number of objects)
+\item A term penalizing deletion of objects (to encourage correct association
+      of objects between frames)
+\item Penalty for inconsistency between actual changes in position (over
+      multi-frame time periods, to avoid noise from single-frame disturbances)
+      and the reported velocity
+\item Penalty for substantial changes in bounding box dimensions (generally,
+      objects stay roughly the same size over time)
+\item Penalty for the amount of area not covered by objects that isn't blue,
+      on the basis that most non-blue things are objects
+\end{itemize}
+
+This did appear to give some meaningful numbers, as when we incorporated
+the HSV water segmentation in with the optical flow algorithm we saw a
+substantial
+improvement in cost (e.g., whereas the cost had been in the range of 125-135
+when tweaking parameters with just optical flow, it cut down to 115 when the
+water segmentation was added). However, we made limited use of the raw
+cost number beyond this.
+
 \section{Results}
@@
 
-\section{Evaluation Section from Proposal}
-
-Because we do not have a labeled data set, we must develop
-some way evaluate the performance of different algorithms.
-
-In order to do this, we propose two general cost functions.
-The first will evaluate the motion tracking features.
-Essentially, when motion tracking, our objective is to
-correctly predict the future path of an object. As such,
-the cost function will assume a sane object detector and
-use the difference between predicted frame-to-frame position
-and actual frame-to-frame position (from the object detector)
-as the cost.
-
-For the object detector itself, a bit more complication is
-required, and we hypothesize that the cost function will
-consist of the following terms:
-
-\begin{description}
-\item[Object Deletion] Any unexpected object deletion shall
-   be penalized, in order to avoid constantly losing track
-   of objects between frames or over/under segmentation of
-   individual objects. ``Expected'' object deletion would
-   occur when an object that is already moving out of frame
-   or behind another object disappears. In theory we also
-   care about excessive object \emph{creation}; however,
-   because the vast majority of created objects are also
-   deleted eventually, the additional complication should
-   be unnecessary. There is an open question about how
-   to treat the deletion of objects that disappear over
-   the horizon; however, such occurrences are hopefully
-   rare enough or easily detectable enough that their
-   cost should be insignificant (we don't want to
-   accidentally \emph{not} penalize flickering objects
-   that happen to fall on the horizon).
-\item[Number of Objects] In order to avoid either a
-   detector that never deletes objects or one that
-   divides real objects into far too many individual
-   objects, we shall penalize the number of objects
-   identified in any given frame. This may be quadratic
-   or some higher order function in order to avoid
-   penalizing small numbers of objects excessively.
-\item[Unaccounted for Area] To avoid an object
-   detector that simply never returns any objects,
-   we shall compare the area accounted for by the
-   identified objects and the area that ``should''
-   contain objects (per some naive estimate), and
-   penalize any unaccounted for area.
-   As an initial method for identifying this area,
-   we propose that we (a) develop some simple detector
-   for identifying sea water (it's mostly blue...),
-   and (b) make a simple assumption about the location
-   of the horizon (the existing video is relatively stable,
-   so the horizon should be, and this way we won't have
-   to worry about the sky). Any area below the estimated
-   horizon that is not sea water ``should'' be part of
-   an object (this includes the shore).
-\end{description}
-
-Obviously, the above cost function will not provide a perfect
-measure of algorithm performance. However, between it and
-qualitative sanity checks, we should be able to provide
-a good estimate of the performance of our algorithms.
-
-\section{Responsibilities Section from Proposal}
-\begin{description}
-\item[James Kuszmaul] Work on basic object detection and how
-  to retain object associations between frames/measure .
-
-\item[Jordan Burklund] Develop methods to segment and mask pixels that represent water. Explore methods using contours for object detection.
-\end{description}
-
 \bibliographystyle{ieeetr}

### 2017-12-14 paper/main.tex
Add diaereses to spellings of naive

@@
 
-\subsection{Naive K-Means Classifier}
+\subsection{Na\"ive K-Means Classifier}
 We initially started with a basic two-class k-means classifier in order to
@@
 to ignore tracking the land anyways. There is a slight bit of noise in the water
-reflections towards the right side of figure \ref{fig:kmeans}, but this naive
+reflections towards the right side of figure \ref{fig:kmeans}, but this na\"ive
 method still provides very useful information.
@@
 \centering
-\caption{Naive K-Means Classifier Results}
+\caption{Na\"ive K-Means Classifier Results}
 \label{fig:kmeans}

### 2017-12-15 paper/main.tex
Minor changes and punctuation

@@
 do not have the processing power to fully take advantage
-of every frame of data, we will pull reasonable samples
-of data to work with.  Specifically, we will generate a test dataset of reduced sized that includes the following important characteristics:
+of every frame of data, we pulled reasonable samples
+of data to work with.  Specifically, we generated a test dataset of reduced sized that includes the following important characteristics:
 
@@
 that the pixels is an object to track.
-The image is converted to HSV color space from the original RGB colorspace since
+The image is converted to HSV color space from the original RGB color space since
 this gives us better segmentation results. The HSV space has better color
@@
 provided the algorithm, as show in figure \ref{fig:kmeans}.  The water pixels
-are mostly labeled as one class, and pixels that not water are mostly labeled as
-another class with a little bit of misclassification noise.  The boat and it's
+are mostly labeled correctly, and pixels that are not water are mostly labeled correctly with a little bit of misclassification noise.  The boat and it's
 wake are clearly labeled as potential obstacles to track, which is what we want.
@@
 boat is upright, and that a majority of the pixels in the bottom half of the
-image are water, and holds true for all of the cases we observed.  If the water
+image are water, which holds true for all of the cases we observed.  If the water
 class ID is not what was expected, the cluster centers are swapped so that the
@@
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=7.8cm]{hsv_kmeans2_orig}
@@
 pixel is to the water class.  The wave reflections observed in figure
-\ref{fig:kmeans} tend to very close to the cluster center while true obstacle
+\ref{fig:kmeans} tend to be very close to the cluster center while true obstacle
 pixels tend to be further away from the cluster center.  By using this measure
@@
 on potential obstacles in the water.  The region of the sky shown in figure
-\ref{fig:kmeans} is classified as a non-water pixel, and we would like to
+\ref{fig:kmeans} is classified as a non-water pixels, and we would like to
 suppress this response to avoid detecting the sky as an obstacle. With only the
@@
 \texttt{bwareaopen} results tended to remove those object to track.  To fix this, an
-erode operation is done on the non-water class pixels to better separate the sky
+erode operation is performed on the non-water class pixels to better separate the sky
 and potential objects to track.
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[height=5cm]{hsv_confidence}
@@
 For the underlying optical flow algorithm, we use Farneback's
-\cite{farneback2003} which is accessible in matlab as
+\cite{farneback2003} which is accessible in Matlab as
 \texttt{opticalFlowFarneback}.
@@
 
-In practice, we just used the default options to the matlab method (which uses
+In practice, we just used the default options to the Matlab method (which uses
 a 3-level pyramid scaling image size by a factor of 2, with a 5-pixel
@@
 \item The camera is an ideal pinhole camera with no substantial lens distortions
-      or the such
 \item There is no sideways relative velocity between the boat and the water
@@
 To actually perform the least-squares fitting we use \texttt{fitlm} in
-matlab. Normally, this just uses a normal least-squares estimation
+Matlab. Normally, this just uses a normal least-squares estimation
 (similar to doing \texttt{A \ b}), but adds some additional
@@
       from the water around it.
-\item The curved metal in the bottom left screws with the optical flow
+\item The curved metal in the bottom left affects the optical flow
       algorithm, causing substantial variation from the model.
@@
 
-As a result of the poor results in the very bottom of the image and in the
+Due to the poor results in the very bottom of the image and in the
 area around the metal bar, we completely ignore and filter out these regions,
@@
 
-We also note that in general objects appearing more distantly in the center of
+We also note that, in general, objects appearing more distantly in the center of
 the image will be hard to identify, as the absolute value of the difference
@@
           object being included, on the assumption that we are interested
-          in keeping tracking going across the longest period of time possible.
+          in maintaining tracking going across the longest period of time possible.
     \item Perform a weighted average of the velocities, weighting by object area.
@@
 \item Around 3 seconds total per frame
+\item Around 200 msec for the HSV segmentation
 \item Around 1 second to call \texttt{opticalFlowFarneback} on the full
@@
 conditions, we had troubles processing some parts of the data. In particular,
-early in certain mornings when not only was the sun interfering with the camera,
+early mornings when not only was the sun interfering with the camera,
 but also in situations such as Figure \ref{fig:mirror}, in which the water is
 relatively flat (reducing texture) and highly reflective, meaning that it is no
-longer particularly blue.
+longer particularly blue. The algorithm also struggles when pointed in the
+ direction of the sun, as the reflections off of the water tend to oversaturate the image sensor.
 
@@
 certainly be pitching and rolling substantially more than the platform we had
-for this footage.
+for this footage, and the weather conditions may be less ideal.
 

### 2017-12-15 paper/two_pager.tex
Add three-page shortened report

--- 2017-12-14 paper/main.tex
@@
 \usepackage{float}
+\usepackage{subfigure}
 \usepackage{chngcntr}
@@
 
-\tableofcontents
-
-\section{Task}
-
 We are currently working on a robotic sailboat for the
-International Robotic Sailing Competition (\url{http://sailbot.org}).
-
-In navigating a sailboat, we would like to be able to
-detect and maneuver around various objects
-that are floating on the surface of the water. As such,
-it is desirable to be able to (a) identify and (b)
-predict the motion of objects on the water. Computer
-Vision is a potentially convenient way to do this.
-We are not specifically aware of any particularly
-thorough published work on this particular application
-(while we have used vision in the past, it has been
-extremely simple detection of bright-colored buoys
-on the water).
-
-The goal of this project shall be to use some existing
-video to try and detect and track objects in video,
-and to understand how well we can develop such a
-detector without having a labelled dataset. We will,
-for the time being, not worry about being able to run
-the algorithm in real time onboard the boat (although
-practical considerations mean it must be able to
-perform fast enough for us to evaluate it).
-
-\section{Data}
-
-We have approximately 240GB of HD video from a GoPro
-mounted on the prow of a large motorboat. While we
-do not have the processing power to fully take advantage
-of every frame of data, we will pull reasonable samples
-of data to work with.  Specifically, we will generate a test dataset of reduced sized that includes the following important characteristics:
-
-\begin{itemize}
-\item With/without land (to see different horizon conditions).
-\item Varying amounts of other objects/traffic.
-\item Areas with significant camera movement in either pitch or roll (may extend to both pitch and roll depending on progress).
-\item Different water conditions (starting with flat, calm water, and progressing to more chaotic water conditions as time allows)
-\item Objects both near/far away (including near the horizon)
-\item Frames that include water droplets on the lens
-\end{itemize}
-
-The boat that the data is collected on is relatively stable
-and the GoPro is mounted in the same spot for the length
-of the video. As such, we should be able to safely ignore
-the rotational pose of the boat (when we apply these
-algorithms to our final boat, we will have to account for
-the much more noticeable movement of that boat).
-\begin{figure}[H]
-\includegraphics[width=12cm]{sample2}
-\centering
-\caption{Sample image 1}
-\end{figure}
+International Robotic Sailing Competition (\url{http://sailbot.org}). For this,
+we have several days of GoPro footage collected from the prow of Professor
+Michalson's motorboat, which we used for developing these algorithms.
 
@@
       based on the content of a single frame, which areas are and
-      are not water/sky/land/objects
+      are not water/sky/land/objects (see Figure \ref{fig:boatconf}:
+  \begin{enumerate}
+    \item We use a nonlinear k-means classification to
+          identify non-water based areas
+    \item And then use Matlab's \texttt{bwareaopen} to
+          identify the connected areas of the sky
+  \end{enumerate}
 \item Use optical flow information from a given pair of frames to
-      locate the moving objects in a frame
-\item Combine the above to determine a set of viable blobs
+      locate the moving objects in a frame, based on Matlab's
+      \texttt{opticalFlowFarneback}, and a simple model of the
+      expected flow of the water to filter out the background in
+      the optical flow space, resulting in confidence scores
+      as in Figure \ref{fig:diff_magnitude}
+\item Combine the above to determine a set of viable blobs using erodes/dilates
+      and Matlab's \texttt{bwlabel}
 \item Associate the blobs with each other and with preexisting objects,
       creating new objects for blobs that can't be associated with any
-      preexisting objects
+      preexisting objects, based on:
+  \begin{enumerate}
+    \item Similarity in blob/object velocities
+    \item Similarity of blob/object position
+    \item Taking weighted averages of velocities to combine them, and
+          combining the object bounding boxes
+  \end{enumerate}
 \item Update objects in preparation for the next frame, updating
@@
 \end{enumerate}
-
-\section{Pixel Level Water Segmentation}
-In the first part of the pipeline, we look at individual images and attempt to
-provide a pixel level detection of potential objects to track.  This step of the
-pipeline provides a confidence value for each pixel in the frame for confidence
-that the pixels is an object to track.
-The image is converted to HSV color space from the original RGB colorspace since
-this gives us better segmentation results. The HSV space has better color
-properties for segmentation since the water is typical near one particular
-color, but can have variations in brightness.  This also allows the segmentation
-to easily segment non-water pixels by the hue value, since the water tends to be
-near the blue and green hue values in our test images.
-
-\subsection{Naive K-Means Classifier}
-We initially started with a basic two-class k-means classifier in order to
-observe how well it would perform.  The classifier was trained on the first
-image in the video sequence, and the L2 distance to the two cluster centers was
-used to classify each pixel in subsequent frames.  Surprisingly enough, the
-2-class k-means classifier worked relatively robustly on the videos that we
-provided the algorithm, as show in figure \ref{fig:kmeans}.  The water pixels
-are mostly labeled as one class, and pixels that not water are mostly labeled as
-another class with a little bit of misclassification noise.  The boat and it's
-wake are clearly labeled as potential obstacles to track, which is what we want.
-The land areas are labeled as water pixels, but for our implementation we want
-to ignore tracking the land anyways. There is a slight bit of noise in the water
-reflections towards the right side of figure \ref{fig:kmeans}, but this naive
-method still provides very useful information.
-
-K-means does not always associate the same class identifier with the water and
-non-water pixels each time, so we must perform an intermediary step to detect
-which class identifier it associates with water pixels.  To do so, we find the
-number of pixels in each class in the bottom half of the image, and use the sums
-to vote for which class is associated with water pixels. This assumes that the
-boat is upright, and that a majority of the pixels in the bottom half of the
-image are water, and holds true for all of the cases we observed.  If the water
-class ID is not what was expected, the cluster centers are swapped so that the
-water class ID is always a known value.
-
-\begin{figure}
-\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
-\includegraphics[width=8.5cm]{hsv_kmeans2_result}
-\centering
-\caption{Naive K-Means Classifier Results}
-\label{fig:kmeans}
-\end{figure}
-
-\subsection{Wave Reflection Suppression}
-Although there is some noise in the classification results, the k-means
-classifier does a decent job of selecting pixels that are areas of water. In
-order to suppress false detections as seen in the right side of the classifier
-results in figure \ref{fig:kmeans}, we look at the distance values between the
-pixel's HSV value and the water pixel cluster center to estimate how similar the
-pixel is to the water class.  The wave reflections observed in figure
-\ref{fig:kmeans} tend to very close to the cluster center while true obstacle
-pixels tend to be further away from the cluster center.  By using this measure
-as a confidence value, we can suppress regions that are more likely to be water,
-and emphasize regions that are more likely to be objects to track.  Using the
-distance metric directly provided decent results, but squaring the distance
-metric value provided significantly better suppression of water and emphasis of
-potential objects.  The squared value was then used as a direct measure of
-confidence that each pixel was a potential object to track.
-
-\subsection{Sky Suppression}
-The k-means classifier provides information about which pixels are water and
-which pixels are not water, but we would also like the algorithm to only focus
-on potential obstacles in the water.  The region of the sky shown in figure
-\ref{fig:kmeans} is classified as a non-water pixel, and we would like to
-suppress this response to avoid detecting the sky as an obstacle. With only the
-wave reflection suppression confidences, most of the sky is suppressed, but
-tends to have higher confidences near land as shown in figure
-\ref{fig:skysuppress}.  Since the sky pixels are all connected, and define a
-very large region, we used the Matlab \texttt{bwareaopen} function that removes binary
-pixels in connected areas that are smaller than a particular size.  By looking
-only at pixels that are classified as "not water", we can detect the sky region
-by filtering for pixel clusters that are larger than 1/4 of the area of the
-image, and set the confidence values for those pixels to zero to suppress them.
-In some cases, potential objects to track were touching the sky region and the
-\texttt{bwareaopen} results tended to remove those object to track.  To fix this, an
-erode operation is done on the non-water class pixels to better separate the sky
-and potential objects to track.
-
-\begin{figure}
-\includegraphics[height=5cm]{hsv_confidence}
-\includegraphics[height=5cm]{hsv_kmeans2_suppressed}
-\centering
-\caption{Sky Suppression Before and After}
-\label{fig:skysuppress}
-\end{figure}
-
-\subsection{Confidence Image Results}
-When the above confidence values and suppressions described above are combined,
-a final confidence value image is produced. Pixels that are more likely to be
-objects to track have higher confidence values than pixels that are more likely
-to be water areas. Figure \ref{fig:boatconf} shows the detector correctly
-detecting a boat in the image.  Figure \ref{fig:towerconf} shows the detector
-correctly detecting a tower in the distance.  Figure \ref{fig:buoyconf} shows
-the detector identifying a buoy off in the distance.
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=7.9cm]{hsv_kmeans2_tower}
-\includegraphics[width=8.3cm]{hsv_kmeans2_tower_result}
-\centering
-\caption{Tower Confidence Results}
-\label{fig:towerconf}
-\end{figure}
-
-\begin{figure}[H]
-\includegraphics[width=7.8cm]{hsv_kmeans2_buoy}
-\includegraphics[width=8.5cm]{hsv_kmeans2_buoy_result}
-\centering
-\caption{Buoy Confidence Results}
-\label{fig:buoyconf}
-\end{figure}
-
-\section{Optical Flow}
-
-\subsection{Underlying Algorithm}
-
-For the underlying optical flow algorithm, we use Farneback's
-\cite{farneback2003} which is accessible in matlab as
-\texttt{opticalFlowFarneback}.
-
-As an overview, the way that this optical flow algorithm works is to
-approximate the neighborhood of every pixel as some quadratic function
-$f(\mathbf{x}) = \mathbf{x}^T \mathbf{A} \mathbf{x} + \mathbf{b}^T \mathbf{x} + c$
-where $\mathbf{x}$ is the location around any given pixel. By estimating these
-polynomials about any given pixel between two frames, we can try to estimate how
-much the image has moved in the neighborhood of that pixel. Actual
-implementations of the algorithm contain some additional filtering/smoothing and
-function on multiple levels of an image pyramid to estimate optical flow at
-different scales. You can also adjust how much of the neighborhood around a
-pixel is used to estimate the polynomial, and various other optimizations.
-
-In practice, we just used the default options to the matlab method (which uses
-a 3-level pyramid scaling image size by a factor of 2, with a 5-pixel
-neighborhood size and a Gaussian filter on the output of size 15x15), as these
-options tended to produce good results.
-
-\subsection{Expected Flow} \label{expected_flow}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.7\textwidth]{example_flow.png}
-\caption{Optical Flow output (shown only for every few pixels) between two
-frames}
-\label{fig:example_flow}
-\end{figure}
-
-In Figure \ref{fig:example_flow} we show the output of
-\texttt{opticalFlowFarneback} on a given pair of frames. The most immediate need
-in this image is to try and account for the field of flow produced along the
-water. This corresponds to the real movement of the plane of the water relative
-to the boat. We can attempt to model this with the following assumptions:
-\begin{itemize}
-\item That the camera is pointed straight and level
-\item There is no substantial roll or yaw angular rates of turn occurring
-\item The camera is an ideal pinhole camera with no substantial lens distortions
-      or the such
-\item There is no sideways relative velocity between the boat and the water
-\end{itemize}
-
-We can define the following variables:
-
-\begin{tabular}{c|c}
-Name & Description \\ \hline
-$h$ & Height of the camera above the water \\
-$v_h$ & The current upwards velocity of the camera \\
-$v_b$ & Forwards velocity of the boat relative to the water \\
-$x_b, y_b$ & Position coordinates in the boat frame \\
-$x_f, y_f$ & X and Y position in the 2-D image frame \\
-$\dot{x}_f, \dot{y}_f$ & The expected optical flow (derivatives of $x_f, y_f$)
-\\
-$K$ & Some constant describing the field of view of the camera \\
-\end{tabular}
-
-The boat frame has an origin immediately below the camera on the surface of
-the water, is attached to the boat, has an x-axis pointing straight forwards,
-a y-axis pointing straight port (left in the camera view) and a z-axis pointing
-straight upwards, with the camera at position $(0, 0, h)$ in this frame.
-
-The image frame is scaled so that the center of the image is the origin and
-the image is exactly 1 unit tall, with the x-axis being the side-to-side axis
-and the y-axis being up-and-down.
-
-Strictly speaking $K = 2\sin\frac\theta2$ if the total vertical field-of-view
-of the camera is $\theta$, but we end up estimating it anyways.
-
-These give us:
-
-\begin{align*}
-x_f =& \frac{-y_b}{Kx_b} \\
-y_f =& \frac{-h}{Kx_b} \\
-\dot{x}_f =& \frac{-y_bv_b}{Kx^2_b} \\
-\dot{y}_f =& -\frac{1}{Kx_b}(\frac{hv_b}{x_b} + v_h) \\
-\end{align*}
-
-And solving for $x_b, y_b$ in terms of $x_f, y_f$ and then substituting in to
-give us the expected optical flow at any given image coordinate $x_f, y_f$ we
-get:
-
-\begin{align}
-\dot{x}_f =& \frac{-v_bKx_fy_f}{h} \\
-\dot{y}_f =& \frac{v_hy_f - Kv_by_f^2}{h}
-\end{align}
-
-If we ignore the actual values of the constants and just turn them into
-something to be estimated, while also adding in a parameter where we
-assume that $y_f$ may not be perfectly centered, we can turn this into a
-linear parameter estimation problem as:
-
-\begin{align}
-\dot{x}_f =& ax_fy_f + bx_f \\
-\dot{y}_f =& cy_f^2 + dy_f + e
-\end{align}
-
-To actually perform the least-squares fitting we use \texttt{fitlm} in
-matlab. Normally, this just uses a normal least-squares estimation
-(similar to doing \texttt{A \ b}), but adds some additional
-options. We tried to make use of these options, including a outlier
-rejection algorithm, but the effects were minimal and substantially
-increased the processing time required to execute the code.
-
 \begin{figure}
@@
 \caption{Difference between the measured and modelled optical flow.
-         Dark blue means perfect match, bright yellow means a poor match.}
+         Dark blue means perfect match, bright yellow means a poor match (and
+         higher likelihood of an object).
+         The sky is later filtered out by ignoring that half of the image.}
 \label{fig:diff_magnitude}
@@
 
-When we compare the expected magnitude of the optical flow to the
-actual magnitude of the optical flow in Figure \ref{fig:diff_magnitude}
-(this is the same image as Figure \ref{fig:example_flow}),
-in which you can see a few features:
-\begin{itemize}
-\item The boat in the left half of the image shows up as a cyan blob, distinct
-      from the water around it.
-\item The curved metal in the bottom left screws with the optical flow
-      algorithm, causing substantial variation from the model.
-\item In the bottom right, due to either lens distortion or hard to measure
-      waves, the optical flow varies substantially from the model.
-\end{itemize}
-
-As a result of the poor results in the very bottom of the image and in the
-area around the metal bar, we completely ignore and filter out these regions,
-as they are relatively unlikely to contain obstacles (or if they did, our
-boat would be running them over anyways).
-
-We also note that in general objects appearing more distantly in the center of
-the image will be hard to identify, as the absolute value of the difference
-in magnitude from the modelled magnitude will be low (because we are looking
-at optical flow in the image, not real life, so far away objects have
-low velocities in the image), and so simple thresholds will not work well.
-We experimented with reducing the required threshold in the center of the image,
-which did improve tracking.
-
-It is also the case that any stationary, low-lying obstacles (e.g. buoys)
-will be indistinguishable from the water and so will not be detected
-by this half of the filtering.
-
-\section{Object Creation and Management}
-
-\subsection{Blob Detection and Filtering}
-
 \begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{threshold_bwlabel}
-\caption{Result of thresholding after dilation/labelling}
-\label{fig:threshold_bwlabel}
+\hfill
+\subfigure[Blobs pre-filtering]{\includegraphics[width=0.3\textwidth]{threshold_bwlabel}}
+\hfill
+\subfigure[Blobs
+post-filtering]{\includegraphics[width=0.3\textwidth]{filtered_labels}}
+\hfill
+\caption{Results of blob detection on HSV and optical flow confidences}
+\label{fig:blobs}
 \end{figure}
-
-From the results in Figures \ref{fig:boatconf} and
-\ref{fig:diff_magnitude}, we have a set of confidence values that can be used to
-identify where individual obstacles may be. In order to combine the two, we
-simply scale each result by some constant and sum the two to provide a
-detection score. Given this score, we then:
-\begin{enumerate}
-\item Threshold against some constant (tunable) value
-\item Do an erode operation using a circular erosion pattern
-\item Do a dilate in reverse, using a radius roughly twice that of the erosion,
-      so that not only is lost space taken up but nearby segments are connected.
-\item Label the connected regions using \texttt{bwlabel} (corresponding with
-      Figure \ref{fig:threshold_bwlabel}).
-\item For each of these regions, throw out any regions that meet any of the
-      following conditions (resulting in Figure \ref{fig:filtered_labels}:
-  \begin{itemize}
-    \item The object is too ``blue'' (by comparing to some hue-based value,
-          in order to try and filter out any water that got detected)
-    \item The object is too small (suggesting it may not really exist)
-    \item The object is too large (sometimes large (~half the image) connected
-          segments will appear that are virtually never real objects)
-    \item It is moving too slowly relative to us (relative to the camera,
-          the only things that should be that constant are on our
-          boat---anything else should at least contain some noise)
-  \end{itemize}
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{filtered_labels}
-\caption{Result of filtering the blobs in Figure \ref{fig:threshold_bwlabel},
-         with the boat and parts of its wake remaining}
-\label{fig:filtered_labels}
-\end{figure}
-
-Following these steps, we initialize each remaining blob as an object, reducing
-it to a bounding box (rather than an exact set of pixels), a median velocity
-(from the optical flow), a position (the center of the bounding box), a label
-(arbitrarily assigned), and a confidence score (initialized to 0.2, to represent
-an object that we are not entirely sure exists yet).
-
-We then go through and, for each object, attempt to both associate it with other
-newly created objects (in case any objects were over-segmented) as well as
-objects that we've kept track of from the previous frame.
-
-\subsection{Object Matching}
-
-When attempting to match objects, we first update all the old objects so that
-their positions have been shifted by their velocities, to better match objects
-that are moving between frames. Furthermore, we reduce the bounding box sizes on
-old objects to reflect a greater uncertainty; if the object really does exist
-over the entire previous bounding box, then that should be reaffirmed by any
-blobs detected in the current frame.
-
-We then perform the following steps:
-
-\begin{enumerate}
-\item For each pair of new objects, as well as each pairing of one new and one
-      old object, we compute a difference score that is the sum of:
-  \begin{itemize}
-  \item The difference in velocity direction
-  \item One minus the quotient in velocity magnitudes (speed), such that the
-        larger of the two speeds shall always be in the denominator
-  \item The difference in the positions (L2-norm) of the two objects, scaled
-        along each axis by the mean width and mean height between the two
-        objects (such that if the objects are perfectly aligned in one dimension
-        and touching edges in the other, they will be a distance of 1.0 apart).
-  \end{itemize}
-\item We threshold every difference, leaving a matrix of object ``neighbors''.
-\item Iterating over the newly created objects (in arbitrary order), we
-      combine that object with it's immediate neighbors, and then remove
-      all the objects that were combined from future
-      consideration\footnote{Ideally, we might do some sort of search where we
-      combine neighbors-of-neighbors and so on, but for simplicities' sake, and
-      to avoid situations where we end up combining large numbers of objects
-      that all are tenuously connected, we just explore one step out from any
-      given object.}.
-\item In order to combine each set of objects that we associated, we:
-  \begin{itemize}
-    \item Create a bounding box that surrounds all of the combined objects
-    \item Give the amalgamation the label (and time-of-existence) of the oldest
-          object being included, on the assumption that we are interested
-          in keeping tracking going across the longest period of time possible.
-    \item Perform a weighted average of the velocities, weighting by object area.
-    \item Take the maximum of the confidence scores for the individual objects,
-          assign it to the result object, and then increment it up slightly on
-          the assumption that any time we are combining more than one object,
-          we are probably becoming more certain in the overall object's
-          existence.
-  \end{itemize}
-\item Any old objects that were not associated with blobs in the current frame
-      have their confidence scores reduced and, if they dip below a confidence
-      of 0, are deleted permanently.
-\item Finally, for every object being tracked, we keep a history of its last 20
-      positions. Once an object accumulates a history of more than around 5
-      frames (to reduce potential for individual frames causing too much noise),
-      we start to compare the object's overall average position change to its
-      velocity. If the two differ by more than a particular threshold, then we
-      reduce the confidence score again. This helps to mitigate object resulting
-      from the wake of the boat, as such objects generally have optical flows
-      inconsistent with where the object actually moves frame-to-frame.
-\end{enumerate}
-
-\section{Objective Evaluation}
-
-While we did not end up making much use of it, we did attempt to write a cost
-function that would offer more objective evaluations of our algorithms (as we do
-not have labeled data to perform some simple check against). The
-motivation for this was both to (a) be able to run our algorithms over larger
-segments of data without manually scoring it and (b) help us to better
-understand our own priorities when writing our detection algorithms.
-
-The cost function ultimately consisted of:
-
-\begin{itemize}
-\item A term penalizing the current number of objects (to avoid bloating the
-      number of objects)
-\item A term penalizing deletion of objects (to encourage correct association
-      of objects between frames)
-\item Penalty for inconsistency between actual changes in position (over
-      multi-frame time periods, to avoid noise from single-frame disturbances)
-      and the reported velocity
-\item Penalty for substantial changes in bounding box dimensions (generally,
-      objects stay roughly the same size over time)
-\item Penalty for the amount of area not covered by objects that isn't blue,
-      on the basis that most non-blue things are objects
-\end{itemize}
-
-This did appear to give some meaningful numbers, as when we incorporated
-the HSV water segmentation in with the optical flow algorithm we saw a
-substantial
-improvement in cost (e.g., whereas the cost had been in the range of 125-135
-when tweaking parameters with just optical flow, it cut down to 115 when the
-water segmentation was added). However, we made limited use of the raw
-cost number beyond this.
 
@@
 \centering
-\includegraphics[width=\textwidth]{example_detection}
+\includegraphics[width=0.5\textwidth]{example_detection}
 \caption{Sample frame of the whole stack running}
@@
 
-\subsection{Timing}
-
-As we are interested in ultimately implementing some of these algorithms on our
-actual robot, we are interested in the run time of our algorithms. These
-are extremely rough numbers, from running only single-threaded on a laptop CPU,
-but were reasonably consistent across runs:
-\begin{itemize}
-\item Around 3 seconds total per frame
-\item Around 1 second to call \texttt{opticalFlowFarneback} on the full
-      1920x1080 image
-\item Around half a second to perform the linear fit described in section
-      \ref{expected_flow}
-\item Around a second and a half spread out among other steps
-\end{itemize}
-
-The linear fit step in particular initially took up a substantial amount of
-time, but we discovered that simply subsampling the data and only using around
-10 percent of it worked almost as well while reducing the runtime to the half
-second seen here.
-
-The optical flow timing is exacerbated by the fact that we are running optical
-flow on the entire, reasonably high-resolution, image. Were we to focus better
-on interest regions, or simply use a lower resolution, we could likely achieve
-substantially better results.
-
-\subsection{Robustness}
-
-\begin{figure}[H]
-\includegraphics[width=12cm]{sample1}
-\centering
-\caption{Mirror-like water}
-\label{fig:mirror}
-\end{figure}
-
-While the algorithms do work reasonably well for footage from these particular
-conditions, we had troubles processing some parts of the data. In particular,
-early in certain mornings when not only was the sun interfering with the camera,
-but also in situations such as Figure \ref{fig:mirror}, in which the water is
-relatively flat (reducing texture) and highly reflective, meaning that it is no
-longer particularly blue.
+For runtime (as we hope to run these algorithms on a boat in real time),
+we currently are looking at around 3 seconds per frame, running single-threaded
+on a laptop CPU. Given reasonable optimizations, lower resolutions, and lower
+frame-rates, we can likely get these times down to work onboard the boat.
 
@@
 
-\bibliographystyle{ieeetr}
-\bibliography{cite}
-
 \end{document}

### 2017-12-15 paper/two_pager.tex
Add back in note on weather

@@
 For future work, we need to actually improve the timing (as mentioned above) to
-work in real time, improve robustness for a greater variety of conditions, and
-to increase the generality to function on a less stable platform--our boat will
+work in real time, improve robustness for a greater variety of conditions
+(including different weather), and
+to increase the generality to function on a less stable platform---our boat will
 certainly be pitching and rolling substantially more than the platform we had

### 2017-12-15 paper/two_pager.tex
Add robustness section back in

@@
-\documentclass{article}
+\documentclass[conference]{IEEEtran}
 \usepackage[utf8]{inputenc}
@@
 \graphicspath{ {images/} }
+
+\begin{document}
+
 \title{RBE/CS549 Project: Boat Tracking/Detection}
-\author{Jordan Burklund \and James Kuszmaul}
-\begin{document}
+\author{\IEEEauthorblockN{Jordan Burklund \and James Kuszmaul}}
 \maketitle
@@
 \begin{figure}[H]
-\includegraphics[width=16cm]{algorithm_flowchart}
+\includegraphics[width=7cm]{algorithm_flowchart}
 \centering
@@
 \begin{figure}[H]
-\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
-\includegraphics[width=8.5cm]{hsv_kmeans2_suppressed}
+\includegraphics[width=7cm]{hsv_kmeans2_orig}
+\includegraphics[width=7.5cm]{hsv_kmeans2_suppressed}
 \centering
@@
 \centering
-\includegraphics[width=0.6\textwidth]{diff_magnitude}
+\includegraphics[width=0.45\textwidth]{diff_magnitude}
 \caption{Difference between the measured and modelled optical flow.
@@
 \hfill
-\subfigure[Blobs pre-filtering]{\includegraphics[width=0.3\textwidth]{threshold_bwlabel}}
+\subfigure[Blobs pre-filtering]{\includegraphics[width=0.23\textwidth]{threshold_bwlabel}}
 \hfill
 \subfigure[Blobs
-post-filtering]{\includegraphics[width=0.3\textwidth]{filtered_labels}}
+post-filtering]{\includegraphics[width=0.23\textwidth]{filtered_labels}}
 \hfill
@@
 \begin{description}
-\item[\url{https://drive.google.com/file/d/1FoKWkp5xDSKwp5YyluKrclo_kGFU4gsg/view?usp=sharing}]
+\sloppy
+\item  \footnote{\url{https://drive.google.com/file/d/1FoKWkp5xDSKwp5YyluKrclo_kGFU4gsg/view?usp=sharing}}
   In this video we have a longer view of the same boat as Figure
   \ref{fig:example_detection}
-\item[\url{https://drive.google.com/file/d/1ndVSwF77_4va6U8rWT4FGJOXq1AjKL_Z/view?usp=sharing}]
+\item \footnote{\url{https://drive.google.com/file/d/1ndVSwF77_4va6U8rWT4FGJOXq1AjKL_Z/view?usp=sharing}}
   This has a slightly shorter capture of a different boat, indicating some
@@
 frame-rates, we can likely get these times down to work onboard the boat.
+
+\section{Robustness}
+
+\begin{figure}[H]
+\includegraphics[width=7cm]{sample1}
+\centering
+\caption{Mirror-like water}
+\label{fig:mirror}
+\end{figure}
+
+While the algorithms do work reasonably well for footage from these particular
+conditions, we had troubles processing some parts of the data. In particular,
+early mornings when not only was the sun interfering with the camera,
+but also in situations such as Figure \ref{fig:mirror}, in which the water is
+relatively flat (reducing texture) and highly reflective, meaning that it is no
+longer particularly blue. The algorithm also struggles when pointed in the
+ direction of the sun, as the reflections off of the water tend to oversaturate the image sensor.
 

### 2017-12-15 paper/two_pager.tex
Add individual contributions
Wording tweaks and margin fix

@@
 \usepackage{graphicx}
-\usepackage[margin=1in]{geometry}
 \usepackage{float}
@@
   \begin{enumerate}
-    \item We use a nonlinear k-means classification to
+    \item Use a 2-class k-means classification to
           identify non-water based areas
-    \item And then use Matlab's \texttt{bwareaopen} to
-          identify the connected areas of the sky
+    \item Suppress wave reflections using squared cluster distance as confidence
+    \item Use Matlab's \texttt{bwareaopen} to
+          identify and suppress the connected areas of the sky
   \end{enumerate}
@@
 \centering
-\caption{Boat Confidence Results}
+\caption{Segmentation Confidence Results}
 \label{fig:boatconf}
@@
 
+\section{Individual Contributions}
+Jordan Burklund
+\begin{itemize}
+
+\item Na\"ive K-means classifier
+\item Water and sky suppression
+\end{itemize}
+
+James Kuszmaul
+\begin{itemize}
+\item Optical flow segmentation
+\item Object blob association
+\item Evaluation cost function
+\end{itemize}
+
 \end{document}

### 2017-12-15 paper/two_pager.tex
Add parentheses

@@
       based on the content of a single frame, which areas are and
-      are not water/sky/land/objects (see Figure \ref{fig:boatconf}:
+      are not water/sky/land/objects (see Figure \ref{fig:boatconf}):
   \begin{enumerate}
